### 2015-03-15 staticanalysis.tex
updates in the static analysis section

The proof sketch behind our static lockset analysis is given in Section~\ref{tech:proof}. The instrumentation and analysis is described in more detail in Section~\ref{tech:method}. The summarisation that we perform to achieve scalability is discussed in Section~\ref{tech:summarisation}. Various optimisations that we perform to achieve precision are described in Section~\ref{tech:optimisations}. Finally, assumption and limitations are discussed in Section~\ref{tech:limitations}.

@@

\subsection{Modelling and Optimisations}
\label{tech:optimisations}

@@

Finally, we also optimise the use of function pointers. LLVM and SMACK do not perform context-sensitive translation, which means that a function that accepts a function pointer as an argument, can arbitrarily call any possible method that corresponds to that function pointer regardless of call site. To increase precision, Chauffeur analyses the AST of the original program and attempts to infer context-sensitive function pointer use information. This information is the used by \whoop to increase precision. To be sound, this optimisation is conservative: if no information can be extracted for a function pointer at a call site, then \whoop uses the original over-approximated function.

\subsection{Assumptions and Limitations}
\label{tech:limitations}

Draft.

>>>>

The proof sketch behind our static lockset analysis is given in Section~\ref{tech:proof}. The instrumentation and analysis is described in more detail in Section~\ref{tech:method}. The summarisation that we perform to achieve scalability is presented in Section~\ref{tech:summarisation}. 
Information about modelling, optimisations and assumptions is given in Section~\ref{tech:optimisations}. Finally, the limitations of our static lockset analysis are discussed in Section~\ref{tech:limitations}.

@@

\subsection{Modelling, Optimisations and Assumptions}
\label{tech:optimisations}

@@

We also optimise the use of function pointers. LLVM and SMACK do not perform context-sensitive translation, which means that a function that accepts a function pointer as an argument, can arbitrarily call any possible method that corresponds to that function pointer regardless of call site. To increase precision, Chauffeur analyses the AST of the original program and attempts to infer context-sensitive function pointer use information. This information is the used by \whoop to increase precision. To be sound, this optimisation is conservative: if no information can be extracted for a function pointer at a call site, then \whoop uses the original over-approximated function.

Finally, we assume that the formal parameters of an entry point do not alias, and thus do not race. This is a potentially unsound feature that can be turned off using a command line option. In our experience so far, though, we have not missed any data races because of this assumption. We also assume that our environmental model and domain-specific knowledge are correct and do not cause any bugs to be missed.

\subsection{Limitations}
\label{tech:limitations}

The main limitation of \whoop is that it is based on static lockset analysis. As an example of this limitation is that we do not check for dynamically created locks (or for locks that are provided by external libraries). Because we perform two thread reduction, we also over-approximate the read accesses, which is another source of false positives. Environmental modelling is also another limitation of \whoop. As aforementioned, enriching \whoop with domain-specific knowledge to increase precision is an ongoing manual effort, but requires expertise in the Linux kernel.

We argue that further increasing the precision of our models is orthogonal to the contributions of this paper. Moreover, despite unavoidable imprecisions during the sound static data race analysis, \whoop can use the reported information to significantly speedup precise bug-finders for concurrent programs, as seen in Sections~\ref{bugfinding} and~\ref{evaluation}.
